---
title: "Survey Results from: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  html_document:
    theme: paper  
    highlight: tango  
    toc: true
    toc_depth: 5
    number_sections: false  
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

pacman::p_load(dplyr,ggplot2)

# loading data
raw <- read.csv("./data/MA_CompPysio_Survey.csv")
metadata <- read.csv("./data/MA_CompPysio_Survey_metadata.csv")

# tidying data
dat <- raw
names(dat) <- metadata$Procesed.Column.Name[match(names(dat), metadata$Raw.Column.Name)]

# tidying categorical options
dat$meta_analysis_self_described <- factor(gsub(" \\(.*", "", dat$meta_analysis_self_described), levels = c("yes","no","unclear/other"))
dat$comparative_physiology_main_focus <- factor(gsub(" \\(.*", "", dat$comparative_physiology_main_focus), levels = c("yes","no","unclear/other"))

# fixing format of journals

#dat$DOI[dat$journal=="2017"] # Typo
dat$journal[dat$journal=="2017"] <- "Biology Letters"

#dat$DOI[grep("Proceedings", dat$journal)] # all the same journal
dat$journal[grep("Proceedings", dat$journal)] <- "Proceedings of the Royal Society B: Biological Sciences"

# standardising DOI format
dat$DOI <- gsub(".*org/", "", dat$DOI)
dat$DOI <-gsub(".* ", "", dat$DOI)

# ggplot2 theme for plots
background.colour <- "white"
text.colour <- "black"
# Font Choice
#loadfonts()
text.font <- "Arial"

tm <- theme(panel.background = element_rect(fill = background.colour),
            plot.background = element_rect(fill = background.colour),
            panel.grid = element_blank(),
            text =  element_text(family = text.font, colour = text.colour),
            axis.text = element_text(colour = text.colour),
            axis.title.x = element_text(margin = margin(10,0,0,0)),
            axis.title.y = element_text(margin = margin(0,10,0,0)),
            axis.line = element_line(colour = text.colour, size = 0.5),
            axis.ticks = element_line(colour = text.colour, size = 0.5),
            plot.title = element_text(hjust = 0.5),
            legend.background = element_rect(fill = background.colour),
            legend.key = element_rect(fill = background.colour),
            plot.margin = margin(6,6,6,6)) 
```

## Sample of `r nrow(dat)` papers {.tabset} 

### Meta-analysis {.tabset .tabset-fade .tabset-pills}   

Answer to the question "Is paper self-describing as containing a meta-analysis (or meta-regression) or actually using meta-analysis/regression approach?".

```{r}
dat %>% dplyr::group_by(meta_analysis_self_described) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = meta_analysis_self_described, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = "meta-analysis or meta-regression",
       y = "number of papers") 
```


### Comparative physiology {.tabset .tabset-fade .tabset-pills}   

Answer to the question ""Is the study's main focus on comparative physiology?"

- yes = focused only on physiological traits very clearly)  
- no =  does have physiological traits, but is interested in broader questions involving many different traits   

```{r}
dat %>% dplyr::group_by(comparative_physiology_main_focus) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = comparative_physiology_main_focus, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = "main focus on comparative physiology",
       y = "number of papers") 
```


### Journals {.tabset .tabset-fade .tabset-pills}   

```{r}
dat %>% dplyr::group_by(journal) %>% dplyr::tally() %>% dplyr::arrange(n) %>% dplyr::mutate(journal = factor(journal, levels = unique(journal))) %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=8)) +
  geom_col(aes(x = journal, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,13,2)) +
  labs(x = "journal",
       y = "number of papers") +
  coord_flip()
```

### Publication years {.tabset .tabset-fade .tabset-pills}   

```{r}
dat %>% dplyr::group_by(pub_year) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = pub_year, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,13,2)) +
  scale_x_continuous(breaks = min(dat$pub_year):max(dat$pub_year)) +
  labs(x = "publication year",
       y = "number of papers")
```



## R Session Information   
```{r}
print(sessionInfo(), locale = F)
```