---
title: "Survey Results from: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Pottier, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  html_document:
    theme: paper  
    highlight: tango  
    toc: true
    toc_depth: 5
    number_sections: false  
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

pacman::p_load(dplyr,ggplot2,here, pander, janitor, magrittr,ggalluvial,ggpubr)

# loading data
raw <- read.csv(here("data","MA_CompPysio_Survey.csv"))
metadata <- read.csv(here("data","MA_CompPysio_Survey_metadata.csv"))

# tidying data
dat <- raw
names(dat) <- metadata$Procesed.Column.Name[match(names(dat), metadata$Raw.Column.Name)]

# tidying categorical options
dat$meta_analysis_self_described <- factor(gsub(" \\(.*", "", dat$meta_analysis_self_described), levels = c("yes","no","unclear/other"))
dat$comparative_physiology_main_focus <- factor(gsub(" \\(.*", "", dat$comparative_physiology_main_focus), levels = c("yes","no","unclear/other"))

# fixing format of journals

#dat$DOI[dat$journal=="2017"] # Typo
dat$journal[dat$journal=="2017"] <- "Biology Letters"

#dat$DOI[grep("Proceedings", dat$journal)] # all the same journal
dat$journal[grep("Proceedings", dat$journal)] <- "Proceedings of the Royal Society B: Biological Sciences"

# standardising DOI format
dat$DOI <- gsub(".*org/", "", dat$DOI)
dat$DOI <-gsub(".* ", "", dat$DOI)

# ggplot2 theme for plots
background.colour <- "white"
text.colour <- "black"
# Font Choice
#loadfonts()
text.font <- "Helvetica"
line.width <- 0.6

tm <- theme(panel.background = element_rect(fill = background.colour),
            plot.background = element_rect(fill = background.colour),
            panel.grid = element_blank(),
            text =  element_text(family = text.font, colour = text.colour),
            axis.text = element_text(colour = text.colour),
            axis.title.x = element_text(margin = margin(10,0,0,0)),
            axis.title.y = element_text(margin = margin(0,10,0,0)),
            axis.line = element_line(colour = text.colour, size = line.width),,
            axis.ticks = element_line(colour = text.colour, size = line.width),
            plot.title = element_text(hjust = line.width),
            legend.background = element_rect(fill = background.colour),
            legend.key = element_rect(fill = background.colour),
            plot.margin = margin(6,6,6,6))

tm2 <- theme(panel.background = element_rect(fill = background.colour),
            plot.background = element_rect(fill = background.colour),
            panel.grid = element_blank(),
            text =  element_text(family = text.font, colour = text.colour, size = 15),
            axis.text = element_text(colour = text.colour),
            axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
            axis.title.x = element_text(margin = margin(10,0,0,0)),
            axis.title.y = element_text(margin = margin(0,10,0,0)),
            axis.line = element_line(colour = text.colour, size = line.width),,
            axis.ticks = element_line(colour = text.colour, size = line.width),
            plot.title = element_text(hjust = line.width),
            legend.background = element_rect(fill = background.colour),
            legend.key = element_rect(fill = background.colour),
            plot.margin = margin(15,15,15,15))
```

## Sample of `r nrow(dat)` papers {.tabset} 

### Meta-analysis {.tabset .tabset-fade .tabset-pills}   

Answer to the question "Is paper self-describing as containing a meta-analysis (or meta-regression) or actually using meta-analysis/regression approach?".

```{r}
dat %>% dplyr::group_by(meta_analysis_self_described) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = meta_analysis_self_described, y = n),
           width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = "meta-analysis or meta-regression",
       y = "number of papers") 
```


### Comparative physiology {.tabset .tabset-fade .tabset-pills}   

Answer to the question ""Is the study's main focus on comparative physiology?"

- yes = focused only on physiological traits very clearly)  
- no =  does have physiological traits, but is interested in broader questions involving many different traits   

```{r}
dat %>% dplyr::group_by(comparative_physiology_main_focus) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = comparative_physiology_main_focus, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = "main focus on comparative physiology",
       y = "number of papers") 
```

How do the two categorisations (meta-analysis and comparative focus) combine?

```{r}
dat %>% 
  ggplot() + tm +
  geom_bar(aes(x = comparative_physiology_main_focus,
               fill = meta_analysis_self_described,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis_d() +
  labs(x = "main focus on comparative physiology",
       y = "number of papers",
       fill = "meta-analysis") 
```


### Effect Sizes Used {.tabset .tabset-fade .tabset-pills}   

```{r}
dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type) %>%
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Means = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=8)) +
  geom_col(aes(x = effect, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,20,2)) +
  labs(x = "Effect Size",
       y = "Number of Papers") +
  coord_flip()
```


### Types of models used {.tabset .tabset-fade .tabset-pills}

What models are used. Note that you need to take this approach because some papers use many different models.

```{r}
summary <-  dat %>% dplyr::group_by(stat_model_type) %>% janitor::tabyl(stat_model_type) %>%
  dplyr::summarise(`Multilevel Meta-analysis Model` = dplyr::if_else(grepl("multilevel meta-analysis", stat_model_type), n, 0),
                  `Multilevel Metaregression Model` = dplyr::if_else(grepl("meta-regression models", stat_model_type), n, 0),
                    `Linear Regression` = dplyr::if_else(grepl("linear regression", stat_model_type), n, 0),
                 Other = dplyr::if_else(grepl("unclear/other", stat_model_type), n, 0),
                 `Vote Counting` = dplyr::if_else(grepl("vote counting", stat_model_type), n, 0),
                 `Weighted Regression` = dplyr::if_else(grepl("weighted regression including mixed effects models", stat_model_type), n, 0),
                 `Random Effects Model` = dplyr::if_else(grepl("random-effects meta-analysis", stat_model_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = (n / sum(n))) %>% arrange(effect)
rownames(summary) <- NULL

summary %>% 
  ggplot() + tm +
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),limits = c(0,0.4)) +
  labs(x = "Analytical Approach",
       y = "Frequency") +
  coord_flip()
```

### Weighted models? {.tabset .tabset-fade .tabset-pills}

Were effect sizes used in the analysis weighted by sampling variance?

```{r}
dat %>%  janitor::tabyl(meta_analysis_weighted) %>% 
  ggplot() + tm +
  geom_col(aes(x = meta_analysis_weighted,  y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(x = "Weighted Model?",
       y = "Frequency") 
```

How do the two categorisations (meta-analysis and weighting) combine?

```{r}
dat %>% 
  ggplot() + tm +
  geom_bar(aes(x = meta_analysis_self_described,
               fill = meta_analysis_weighted,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_fill_viridis_d() +
  labs(x = "Meta-analysis?",
       y = "Proportion",
       fill = "Weighted?") 
```

### Effect size impacted by 'nuisance heterogeneity'? {.tabset .tabset-fade .tabset-pills}

Could effect size be impacted by an additional continous predictor (e.g., temperature in thermal limit studies, dosage in hormone manipulation studies)?

```{r}
dat %>%  janitor::tabyl(cont_moderator_variable) %>% 
  ggplot() + tm +
  geom_col(aes(x = cont_moderator_variable, y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(x = "Nuisance Heterogenity?",
       y = "Frequency") 
```

### Dealt with nuisance in model? {.tabset .tabset-fade .tabset-pills}

Of the studies that said 'yes' how many actually modelled it or accounted for it?

```{r}
dat %>%  filter(!cont_moderator_variable_modelled=="") %>% janitor::tabyl(cont_moderator_variable_modelled) %>% 
  ggplot() + tm +
  geom_col(aes(x = cont_moderator_variable_modelled, y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(x = "Nuisance Heterogenity Modelled?",
       y = "Frequency") 
```

```{r}
dat %>% mutate(cont_moderator_variable_modelled = if_else(cont_moderator_variable_modelled == "", "N/A", cont_moderator_variable_modelled)) %>% 
  ggplot() + tm +
  geom_bar(aes(x = cont_moderator_variable,
               fill = cont_moderator_variable_modelled,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_fill_viridis_d() +
  labs(x = "Nuisance Heterogenity?",
       y = "Proportion",
       fill = "Modelled Nuisance\n Heterogenity?") 
```

### Reported Heterogeneity? {.tabset .tabset-fade .tabset-pills}

Of the studies that are weighted how many actually reported measures of heterogeinity?

```{r}
dat %>%  filter(!heterogeneity_reported=="N/A (e.g. they conducted unweighted meta-analysis)") %>% mutate(heterogeneity_reported = if_else(heterogeneity_reported == "unclear/other (add comment)", "other", heterogeneity_reported)) %>% janitor::tabyl(heterogeneity_reported) %>% 
  ggplot() + tm +
  geom_col(aes(x = heterogeneity_reported, y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(x = "Heterogeneity reported?",
       y = "Frequency") 
```

### Publication years {.tabset .tabset-fade .tabset-pills}   

```{r}
dat %>% dplyr::group_by(pub_year) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = pub_year, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,13,2)) +
  scale_x_continuous(breaks = min(dat$pub_year):max(dat$pub_year)) +
  labs(x = "publication year",
       y = "number of papers")
```


### Handling sources of non-independence {.tabset .tabset-fade .tabset-pills}

Did the analysis account for - or mention - additional non-independence arising from, e.g., phylogenetic relatedness or shared control treatments? The actual way of dealing with non-independence (n-i) will vary from study to study.

```{r}
dat %>%  janitor::tabyl(nonindependence_mentioned) %>% 
  ggplot() + tm +
  geom_col(aes(x = nonindependence_mentioned, y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_x_discrete(labels = c('n-i dealt with w/o modelling', 'n-i modelled', 'unclear/other/not mentioned')) +
  labs(x = "Non-independence handling",
       y = "Frequency") 
```

### Publication bias {.tabset .tabset-fade .tabset-pills}

What type of publication bias analysis was used in the analysis? Multiple metrics can be used per paper, hence frequencies won't sum to 100%.

```{r}
dat %>% dplyr::group_by(pub_bias_type) %>% janitor::tabyl(pub_bias_type) %>%
  dplyr::summarise(GraphicalTest = dplyr::if_else(grepl("graphical tests", pub_bias_type), n, 0),
                  RegressionTest = dplyr::if_else(grepl("regression test", pub_bias_type), n, 0),
                    TrimAndFill = dplyr::if_else(grepl("trim-and-fill", pub_bias_type), n, 0),
                 TimeLag = dplyr::if_else(grepl("time-lag", pub_bias_type), n, 0),
                 OtherUnclear = dplyr::if_else(grepl("unclear", pub_bias_type), n, 0),
                 None = dplyr::if_else(grepl("no", pub_bias_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = (n / sum(n)*100)) %>% arrange(effect) %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=8)) +
  geom_col(aes(x = effect, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,30,2)) +
  scale_x_discrete(limits = c("None", "GraphicalTest", "RegressionTest", "TrimAndFill",
                              "TimeLag", "OtherUnclear")) +
  labs(x = "Effect Size",
       y = "Number of Papers") +
  coord_flip()
```


### Synthetic plot

```{r}
plot1 <- dat %>% 
  ggplot() + tm2 + theme(legend.position = "top") +
  geom_bar(aes(x = comparative_physiology_main_focus,
               fill = cont_moderator_variable,
               y = ..count../sum(..count..)), width = 0.5,
           colour = NA, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Main focus on comparative physiology",
       y = "Number of papers",
       fill = "Continuous predictor present")
plot1

plot2 <- dat %>% 
  ggplot() + tm2 + theme(legend.position = "top") +
  geom_bar(aes(x = pub_year,
               fill = meta_analysis_self_described,
               y = ..count../sum(..count..)), width = 0.5,
           colour = NA, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Year of publication",
       y = "Number of papers",
       fill = "Meta-analysis")
plot2

plot3 <- dat %>%
  ggplot(mapping = aes(x = comparative_physiology_main_focus,
               y = meta_analysis_self_described)) + tm2 +
  theme(panel.border = element_rect(size = 0.5, fill = NA),
        panel.grid.major = element_line(colour = "grey92"),
        axis.text.y = element_text(angle = 90)) +
  geom_count(mapping = aes(size = ..n.., colour = "A")) +
  scale_size_continuous(range = c(20,40), guide = F) +
  scale_colour_brewer(palette = "Set2", guide = F) +
  # scale_y_continuous(expand=c(0,0)) +
  # scale_fill_viridis_d() +
  labs(x = "Main focus on comparative physiology",
       y = "Number of papers")
plot3

plot4 <- dat %>%
  ggplot(aes(axis1 = meta_analysis_self_described,
             axis2 = comparative_physiology_main_focus,
             axis3 = cont_moderator_variable)) +
  ggalluvial::geom_alluvium(aes(fill = comparative_physiology_main_focus), width = 1/4, alpha = 0.8) +
  tm2 + theme(axis.line = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none", panel.grid = element_blank(), axis.text.y = element_blank()) +
  scale_fill_brewer(palette = "Set2") +
  ggalluvial::geom_stratum(fill = 'white', colour = 'black', width = 1/4) +
  geom_text(stat = 'stratum',
            aes(label = after_stat(stratum)),
            # label = c('weighted', 'unweighted', 'unclear'),
            colour = 'black') +
  scale_x_discrete(expand = c(.1, .1), limits = c("Meta-analysis", "Comparative physiol.", "Cont. moderator"))
plot4


panels <- ggarrange(ggarrange(plot1, plot2, nrow = 1),
                    ggarrange(plot3, plot4, nrow = 1, widths = c(1,2)),
                    nrow = 2, ncol = 1)
ggsave(panels, filename = "multipanel.pdf", device = "pdf")

```


## R Session Information   
```{r}
pander(sessionInfo(), locale = FALSE)
```