---
title: "Meta-analytic approaches and 'new' effect sizes to account for heterogeneous effects in comparative physiology"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: no
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
  ## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
    ## and rounded to 2 digits		  ## and rounded to 2 digits
    options(digits = 2)
    
  pacman::p_load(flextable, tidyverse, equatags)  
```

Daniel W.A. Noble$^{1,*}$, Patrice Potter$^{2}$, Sammy Burke$^{2}$, Szymek Drobniak$^{2}$, Malgorzata Lagisz$^{2}$, Rose O'Dea$^{2}$, Shinichi Nakagawa$^{2}$

$^{1}$ Division of Ecology and Evolution, Research School of Biology, The Australian National University, Canberra, ACT 2600

$^{2}$ School of Biological, Earth and Environmental Sciences, University of New South Wales, Sydney, NSW

## **Abstract**

Meta-analysis is a tool that provides comparative physiologists with powerful, unbiased and quantitatively informed answers to topical research questions by synthesising effect sizes across disparate studies. Meta-analysis's core objective and strength is to distill published research results into standardised effect sizes that can be weighted (i.e., by their inverse sampling variance), and which can be compared across broad sets of research designs, study systems and species. Estimating overall effect sizes, and understanding what drives effect variability, provides opportunities to model how organisms will respond to pressing global challenges. Despite this ambition, research designs in comparative physiology can appear, at the outset, as being vastly different to each other (e.g., using different temperatures or treatment dosages). Differences in treatments across studies has led many to believe that meta-analysis is an exercise in comparing "apples with oranges". Here, we dispel this myth by showing how standardised effects sizes can be used in conjunction with powerful multi-level meta-regression models to both account for factors driving differences across studies and make them more comparable. In addition, we derive new effect size measures that provide comparative physiologists with a means to directly make effect measures comparable without the need to resort to more complex statistical models that may be harder to interpret. Our 'new' effect sizes and corresponding sampling variances help physiologists deal with common treatment differeces across studies; allowing researchers to compare both mean differences and associated differences in variance for physiological traits. The new effect sizes we propose, in combination with existing meta-analytic models, will pave the way for comparative physiologists to explore exciting new questions by making results from large-scale data sets collected from the literature more accessible and widely interpretable.

<!--- Dan - new effect sizes here is a big vague. Don't we just say - we formalize Q10 for by deriving the sampling error variance (the inverse of weight) of its natural logarithm Further, we propose two additional related effect sizes, which use Q10's standard deviation and coefficence of variation (CV). Also, this is an examplar of what eco-phsiologists are familiar with can be used as a standaridsed effect size and they can draw on the tools developed for meta-analysis --->

<!--- xxxx --->

## **Introduction**

A meta-analysis, being part of systematic review, provides quantitatively informed answers to research questions by using a transparent and reproducible search strategy to synthesise research results across studies and identify sources of variation [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Arnqvist1995; @Koricheva2013]. <!--- note meta-analysis itself is not really about transparency etc - it is a stasitical tool - these other bits come in as a part of systeamtic review ---> Meta-analysing independent studies provides greater statistical power and precision than what any individual study on its own would be able to provide on a question - particularly given that most empirical studies are already under-powered in many areas of biology [@Forstmeier2017; @Jennions2003; @Button2013]. By expressing study effects on a common scale (i.e., standardised effect size) a meta-analyst can gain broader insight into the direction and efficacy of a particular treatment effect or the strength of a relationship between two variables of interest [@Gurevitch2018; @Koricheva2013]. Quantitative syntheses, including meta-analyses, have provided comparative physiologists with powerful insights on pressing global challenges - from testing whether physiological plasticity can buffer organisms against climate change [e.g, @Seebacher2015] to the degree to which endocrine disrupting chemicals, such as bisphenol A (BPA), impact aquatic organisms [e.g., @WuSeebacher2020].

Meta-analysis is a well-developed research topic with a long history in statistics, medicine, psychology and ecology [@Gurevitch2018; @Nakagawa2017; @Glass2015]. It has emerged as the 'gold-standard' across various disciplines for quantitative synthesis and is becoming increasingly more prominent in the field of comparative physiology. Formal meta-analyses, as a part of a systematic review, involve a systematic literature-based search, extraction of data from studies meeting specific inclusion criteria, and the subsequent calculation of effects sizes from summary statistics provided by papers. The entire process should follow and report upon this entire process carefully to ensure the meta-analysis within a systematic review, is as transparent and reproducible as possible [i.e., see the Preferred Reporting Items for Systematic Reviews and Meta-Analyses in Ecology and Evolution (PRISMA-EcoEvo) protocol, @ODea2021]. Aside from providing an up-to-date overview of the research question, along with identifying any gaps and biases, there are two major goals of meta-analysis: 1) to provide an overall mean estimate of a treatment effect or relationship and 2) to quantify effect size variance and understand key drivers explaining differences in effects across studies [@Nakagawa2017; @Gurevitch1999; @Gurevitch2018; @Cooper2009; @Koricheva2013]. A critically important feature of meta-analaysis is its ability to account for sampling error [@Koricheva2013]. The inverse of sampling error variance is used to weight studies, accordingly, in the analysis and make overall estimates more precise. While many meta-analyses might employ unweighted analyses, these studies will not gain greater precision on estimates because sampling variance cannot be formally accounted for in the analysis, even if the overall mean itself may be unbiased. <!--- cite Morrissey here - also, the overall mean is unbiased given there is no publication bias or small study effects - this is important as many meta-analyses are likley to suffer from small study effects; my pick sann does not seem like you have talked about publcaiton bias too much - an ablity of meta-analysis to detect publicaiton bias is one of its strength - or probalby the major strength...  --->

Despite the widespread adoption and well established methodological procedures of meta-analysis, it is often criticised for mixing 'apples and oranges' -- in other words, mixing effects from studies that lack comparability [@Gurevitch2018; @Arnqvist1995; @Gallo1979; @Carpenter2020; @Stewart2010; @Glass2015]. Lack of comparability could result from studies differing in experimental design, species, measurement variables and sampling units, which can impact the generalisability of meta-analytic results [@Arnqvist1995; @Stewart2010]. Indeed, study specific differences in comparative physiology are widespread. For example, experimental designs can vary in the temperatures that they apply or in the dosages of chemical modifiers used in experimental treatments [e.g., @WuSeebacher2020]. In addition, they often include a diversity of study species which can vary in responses. To the uninitiated, the concern might not be unfounded given that meta-analyses in ecology and evolution often mix highly heterogeneous effect size estimates, resulting in high variability [@Senior2016]. <!--- this last sentence does not seem to fit well or not completely logical - did you want to say -heterogeniety is normal. Actually Jessica always say in ecology meta-analysis is probalby not that useful; meta-regression is what we are most interested - I think yo cna make a similar point or you are already making one but make it clearer  - the first sentence is a good one but rather than focusing on "contolling" for hetero but finding out what causes hetero should our (comparative phsiologists') main aim for meta-analysis. I know you are probably intending to say something like this but not coming through at the moment. Also, I note that "review generated evidence" only comes from meta-regression not meta-analysis = intercept models - as study general evidnece can be equivalent to meta-analytic results --->

The goal of our review is to briefly overview meta-analytic practices and reassure comparative physiologists that heterogeneity among effect sizes can be comfortably dealt with by modifying existing effect sizes measures and using multilevel meta-regression models. We first conduct a survey of published meta-analyses in the field of comparative physiology to gauge methodological and meta-analytic approaches - assessing what types of effect sizes are commonly used, how often effect size estimates are impacted by continuous treatment differences and whether meta-analytic models (i.e., models that account for sampling variance) are commonly applied. We then introduce new effect sizes and their associated sampling variance to provide comparative physiologists opportunities to explicitly incorporate treatment differences into the effect size estimates to ease their interpretation. <!--- see my comments above - they are not really new effec sizes - just formalising we can chat about this lnRR, lnVR and lnCVR and Q10 all exisit already ---> We then describe how more complex treatment differences, such as dosage differences, which may have non-linear effects, can be accommodated using multilevel meta-regression models. We hope to show how expanding the meta-analytic toolkit for comparative physiologists will provide new opportunities to address global scale problems that attempt to understand how organisms will cope with rapidly changing environments in the future.

## **The Apples and Oranges 'Problem' in Comparative Physiology**

<!---  Dan syas: I'm thinking now, in re-reading this, that we could use a slightly expanded version of paragraph 4 taht describes this problem in a little more detail. This is likely the place for a figure, which I am having some trouble envisaging where one is best placed. It just otherwise seems like a bit of a jump to otherwise head straight into the survey. Here we should also talk about quality control, the need to be transparent and have careful reporting. This is a big issue in medicine. Talked about a lot. Integrate blurb below…try to take a balanced approach. I will add an introduction bit to this paragraph below to put into context…. xxxx --->

In the words of Gene Glass, the father of meta-analysis: "Of course it mixes apples and oranges; in the study of fruit nothing else is sensible; comparing apples and oranges is the only endeavor worthy of true scientists; comparing apples to apples is trivial" [p.g., 224, @Glass2015]. @Glass2015's intention was to simply highlight that whether comparing apples and oranges is a good idea simply boils down to the specific question of interest and the population one wishes to make conclusions about. If we are interested in gaining insight into 'fruit' then synthesising apples and oranges is the only thing that makes sense. <!---  as said, comparative analysis by definition, compares applies and oranges - = different species - I guess you are mainly talking about mixing different methods ones - this is where I said earlier - focusing on explaining hetero comes in. Also, mixing different quality of studies are often referred to as "Garbage in and garbage out" ---> Instead, if we are interested in generalising only to apples then mixing fruit will not be ideal. We tend to agree with @Glass2015, and his sentiment is of particular relevance to comparative physiology. For example, if we want to understand the impact of a pollutant on reproduction in aquatic organisms then including aquatic mammals and fish would provide answers to this general question, even if their reproductive biology and physiology are known to differ. By synthesizing diverse study outcomes, it allows us to not just report on overall effects (if the question is relevant to ask), but understand critical features of the literature, study systems (e.g., different effects of pollutants on fish and mammal reproduction) and approaches that can explain the diversity of effect outcomes observed. In doing so, it provides a richer set of conclusions that can draw attention to important sources of variability that might have otherwise been overlooked. Having said that, it is still very important to understand the limitations and diversity of studies included within a meta-analysis - an important reason why transparency and reproducibility are such a prominent feature of meta-analysis (or ideally should be) [@ODea2021]. The validity of general conclusions to certain questions (e.g., what is the overall effect of a polluntant on reproduction?) will inevitably depend on the homogeneity of effects being synthesized. <!--- this is know as putting one number on so many different kinds of studies - but of course, this may not be that interesting apart from a starting point for subsequent meta-regressoin analyses ---> For example, given the very different physiology of aquatic mammals and fish, understanding an overall combined effect on reproduction may not be particularly relevant, and indeed may be misleading. It is important to consider these questions when conducting meta-analyses.

Of course, there are certain types of heterogeneity that transcend different questions. In other words, if we compare different types of 'apples' is it still fair to synthesize these studies to understand overall effects? This situation might be analogous to a meta-analysis in comparative physiology that attempts to understand the effects of temperature on fish growth. <!--- taxnomically speaking - comparing different applies are comparing different populations of the same spieces of fish so not quite sure whether this is quite correct - I am being of course a bit pedentic---> Studies might rear populations of fish under different temperatures, but in the end, they are still all fish, governed by the same thermodynamic effects on growth. In this situation, we might still be interested in understanding overall effects, but we may question whether such a conclusion is sensible given the diversity of temperatures applied. Such a study does not preclude us from also exploring important moderators or predictors of variability in effects across studies.....

## **Survey of Meta-Analytic Practices in Comparative Physiology**

As discussed above, it is clear that different types of questions and experimental approaches will be more or less susceptable to the 'apples and oranges' problem. <!--- not sure whehter it is a good idea to hightlight it as a problem as said earlier - explaining hetero is what's interesting and even the main porupose of meta-analysis for comparative physiologiests - they are not the problem but they are what we are going to invesitgate. I do not think 'suspetible' is a good word here given the average hetero = > 90% - you are nearly 100% guranteed to have hetero and this is what's interesting ---> As such, before embarking on our path to 'meta-analytic enlightenment', we conducted a literatrure survey to understand how widespread the problem may be in comparative physiology and gauge the ways in which it is currently being addressed, if at all. To accomplish this goal, we searched for meta-analyses and comparative analyses that tackled questions using physiological data compiled from animal studies across multiple species. We were specifically interested in asking: 1) what type of effect size was used within the quantitative synthesis? 2) based on the methods provided, was the effect size likely susceptable to variation in how treatments are applied across studies? 3) If so, was the variation resulting from treatment differences accounted for in the effect size or during analysis?, and finally, 4) how were effect sizes analysed in the quantitative synthesis, were these weighted analyses, and what program was used for analysis? We also recorded whether publication bias and sensitivity analyses were conducted, if data were publicly available, where it was stored, and if so, if it was usable (i.e., in a machine readable format).

#### *Literature Search and Survey Data Collection*

All searches were conducted in Scopus between May 16 - 18, 2021. Briefly, we first did a series of pilot searches (detailed in Supplementary Materials) to refine our search query. We searched for studies that included: 'meta-analysis', 'meta regression' , 'comparative analysis', 'comprehensive analysis' 'global analysis' or 'macrophysiology' in their title and abstract, restricting our search to key comparative physiology journals (deatiled in Table X - Supplementary Materials) within the last 6 years. We expected this to provide a contemporay overview of quantitative syntheses in comparative physiology while making the survey and screening feasible. We awknowledge that this may have missed some important, and influential, quantitative synthese in the field of comparative physiology, but likely provided a random sample to inform on the questions we sought to address (see below).

We included quantitative syntheses, including meta-analyses (*sensu stricto*), that were focused on comparative physiologcial questions only. Narrative reviews and comments were not considered because they did not collate quantitative data from the literature. Studies needed to have collected physiological trait data from the literature including traits, such as hormones, thermal physiology (CTmax, CTmin, thermal preferences, thermal breadth, LT50, lethal thresholds), metabolic rates (routine, maximum, and minimum metabolic rate, aerobic scope), enzyme reaction rates and concentrations, oxidative stress and consequences (e.g., DNA damage, teleomere attrition), performance measurements (e.g., heart rates, sprint speed, swim speed, growth rates), body composition (e.g., fatty acid profiles) and immune traits (e.g., cell function, antibodies, ect.). Studies were also included if they measured the relationship between physiological traits and body size (e.g., metabolic scaling), reproductive allocation, or survival / mortality across taxa. Studies on behavior were only included if the synthesis was explicitly interested in the proximate physiological drivers of behavioral variation. We also included meta-analyses that quantified plasticity in physiological traits across taxa for the purpose of understanding adaptive plasticity in response to environmental variation. While many studies in comparative physiology collect data as part of a comparative analysis, we focus here only on studies collating data from existing literature because this is the major data source for quantitative synthesis. Quantitative syntheses interested in phenology, biomechanics, sexual size dimorphism (i.e., SSD, except if the study was interested in sexual differences in physiological traits), those focused on validating physiological methodology, or characterising population genetic structure, gene transcription and quantitative genetics were excluded. There were also meta-analyses that focused on colour evolution and contrast which varied in their goals but were most often not focused on physiological questions. As such, these studies were also exclude. We restricted our focus to animal studies because plant synthesis papers were often only focused on ecological (e.g., quantified changes in abundance and distribution of species) or agricultural questions (e.g., studies on soil microbial function, nitrogen and phosphorus capture, methane output and greenhouse gas emissions).

All title, abstract and full text screening was done by two people (DWAN and ML). Any conflicts or disagreements were resolved by discussion and a consensus on a paper was reached. Out of the 417 papers, 79 full texts were screened and we included a total of 62 for final data extraction. For full details on the search strings used, dates searched, the number of papers found, and our PRISMA flowchart (along with reasons for exclusion at the full text stage) we refer readers to the *Supplementary Materials*.

#### *Current Meta-analytic Practices in Comparative Physiology*

## **Opening Up New Opportunities in Comparative Physiology: Expanding the Breadth of Effect Sizes to Deal with Treatment Differences Across Studies**

#### *Brief Introduction to Common Effect Sizes and Sampling Errors that are Useful in Comparative Physiology*

Meta-analysis requires comparative physiologists to make use of effect sizes that are suitable for answering their specific question. The key features of all effect sizes used in meta-analyses are that they are: 1) statistical parameters that have been placed on a common scale so that they can be compared across studies and 2) that they have some associated measure of precision (i.e., sampling variance)[@Koricheva2013; @Borenstein2009; @Schmid2021]. Effect sizes in comparative physiology are quite variable; they can be the mean difference between two groups (e.g., difference in mean between a control and treatment group - see @WuSeebacher2020), the slope of a regression (e.g., between body size and metabolic rate), a correlation between two variables, or even the raw mean (or variance) itself (e.g., comparing CTmax or CTmin across species - see @Sunday2014). In many cases, multiple types of effect sizes might be used depending on the question. For example, studies interested in the effect of temperature on some physiological trait that have measured the trait across studies experimentally manipulating temperature could analyse the mean 'effect' by creating contrast-based effect sizes or by analysing the mean in each temperature treatment. The choice as to which one to use will boil down to the types of data available from primary studies, the specific question of interest and whether studies are observational or experimental in nature. Considering our temperature example above, if we are interested in the magnitude of change between temperature treatments, we may want to use a contrast-based effect size, such as the log response ratio $\ln{RR}$. If we are instead interested in how the mean of some physiological trait changes with tempearture we may instead model the mean itself. In the latter scenerio, so long as all the means are measured in the same units (e.g., ; or can be converted to the same units) then the mean may provide the desired answer to a given question. We provide details on different effect measures and associated sample sizes in Table \@ref(tab:tablea).

```{r tablea, echo = FALSE, tab.cap = "Common effect sizes, associated sampling variances and notes on when they might be used. Note that for log odds ratio a continuity correction is applied (i.e., 0.5) to each proportion. In addition, a correction factor, J (1-3 ), is applied to the standardised to correct for small sample effects"}

# I'm just listing a bunch here, but we should think about what we want to include. Here's an example of how we can make a table with all equations. Fun! Really useful website here: https://math.meta.stackexchange.com/questions/5020/mathjax-basic-tutorial-and-quick-reference. Something wrong with character coding when rendering word document. Has to do with the log odds ratio equation. Renders fine in html, but just stops when word
names <- c("Mean",
           "Log Standard deviation. lnSD",
           "Log Response Ratio, lnRR",
           # leave this for now??? - we introduce this later
           "Standardised Mean Difference, SMD",
           "Zr (Fisher Trasnformaiton of \n correlation Coefficient, r)")
           #"Log Odds Ratio",
           #"")

# the index i is not really applied consistently so I am adding it more (to do)
# we shoudl discuss what kinds of notations we want to use consistely
           
eqns_yi <- c("M_i",
             "\\ln{SD_i} + \\frac{1}{2(N_i - 1)}",
             # Senior will have a better version but this is good for now
             "\\ln \\left( \\frac{M_{i1}}{M_{i2}}\\right)",
             "\\frac{\\left( X_{i2} - X_{i1}\\right)}{SD_{p}}J",
             # we need to spell out SD_p somewhere
             "\\frac{1}{2}log\\left(\\frac{1+r_{i}}{1-r_{i}}\\right)")
             #"log\\left(\\frac{y_{i2}+0.5}{n_{i2}-y_{i2}+0.5}\\right) ",
             #"- log\\left(\\frac{y_{i1}+0.5}{n_{i1}-y_{i1}+0.5}\\right)")
             
# Just placeholders for now
eqns_vi <- c("\\frac{SD_{i}^2}{N_{i}}",
             "\\frac{1}{2(N_i - 1)}",
             "\\frac{N_{1i} + N_{2i}}{N_{1i}N_{2i}} + \\frac{SMD^2}{2(N_{1i} + N_{2i})}",
             "\\frac{SD_{1i}} {M^2_{1i}N_{1i}} + \\frac{SD_{2i}} {M^2_{2i}N_{2i}}",
             # see my commenst
             "\\frac{1}{N_i - 3}")
             #"\\frac{s_{i}^2}{n_{i}}",
             #"")
         

table_fin <- data.frame(names, eqns_yi, eqns_vi)

# Note: It is esssential that you be clear that the function 'compose' is exported from the flextable package. That's because purr also has the same function, so it gets messy.
tablea <- flextable(table_fin)                                                                                %>% 
          flextable::compose(j = "names", part = "header", value = as_paragraph(as_b("Effect Measure")))      %>% 
          flextable::compose(j = "eqns_yi", part = "header", value = as_paragraph(as_b("Definition")))        %>% 
          flextable::compose(j = "eqns_vi", part = "header", value = as_paragraph(as_b("Sampling Variance"))) %>%
          flextable::compose(j = "eqns_yi", value = as_paragraph(as_equation(eqns_yi)))                       %>%
          flextable::compose(j = "eqns_vi", value = as_paragraph(as_equation(eqns_vi)))                       %>% 
          flextable::width(j = 1, 3)                                                                          %>% 
          flextable::width(j = 2, 2)

tablea

```

#### *Expanding the Scope of Effect Size Metrics in Comparative Physiology*

A very common occurance in comparative physiology is to manipulate the tempearture organisms experience and measure some physiologcal rate (e.g., metabolic rate). Using normal effect size measures (see above) would result in the effect sizes varying in the types of temperatures applied within and across across studies contributing to 'comparablity' issues. As such, it is often of interest to standardsise effects by temperature. One common effect size measure already used in comparative physiology to compare physiological rates is $Q_{10}$ [e.g., @Rodgers2021; @Seebacher2015; @Havird2020], which describes the multiplicative change in physiological rates across a 10°C temperature change. Higher $Q_{10}$ values indicate larger changes in physiological rates. Currently, however, we do not have associated sampling variances derived for $Q_{10}$ making it challenging to make use of the full power of weighted meta-analytic models. Interestingly, $Q_{10}$ can be seen as a variant of *lnRR*, meaning that we can derive a $Q_{10}$ based effect size and sampling error using the well known mathematical properties of *lnRR*. This opens up the meta-analytic toolbox and improves upon our ability to account for well known sources of non-independence that are typical in meta-analysis [@Lajeunesse2011; @Noble2017].

#### *Comparing changes in mean physiological rates*

Prior to showing how the relevant $Q_{10}$ effect size can be calculated it can be helpful to understand its similarities to *lnRR*. The *lnRR* described by @Hedges1999 and extended by @Lajeunesse2011 and can be calculated as follows <!---put see also Senior et al. 2020 RSM as well - there we expand the 2nd order delta method stuff - so unbiased formlas for all lnRR, lnVR etc is a bit different - we do not want to put it here but mention by the way - Lajeuness 2011 is not the one I would cite Lajeuness 2015 is more relevant ---> :

$$
\begin{equation} 
  \begin{aligned}
  lnRR = ln\left ( \frac{X_{1}}{X_{2}}\right )
  \end{aligned}
  (\#eq:lnRR)
\end{equation} 
$$

$$
\begin{equation} 
  s_{lnRR} = \left ( \frac{SD_{1}^2}{N_{1}X_{1}^2}\right ) + \left ( \frac{SD_{2}^2}{N_{2}X_{2}^2}\right )
  (\#eq:slnRR)
\end{equation} 
$$ In equation \@ref(eq:lnRR), $X_{1}$ is the mean of group 1, often a control group, where as $X_{2}$ is the mean of group 2 (e.g., a treatment group). The mean of $X_{i}$ for group *i* can be any measurement type (e.g. a physiological rate, mass etc) so long as the measurement variable is ratio scale. Log transformation of this ratio makes the effect size normally distributed. <!--- use consistent notations in relation to the table - do you want to put i or not - then you need to put i throughout ---> Equation \@ref(eq:slnRR) is the analytical solution for *lnRR*'s sampling variance where, $SD_{1}^2$ and $SD_{2}^2$ is the sample variance in group 1 and 2, respectively and $N_{1}$ and $N_{2}$ is the sample size in group 1 and 2.

The equations for *lnRR* and its sampling variance allow us to easily extend this to $Q_{10}$ based effect sizes. Recall that $Q_{10}$ is described as follows:

$$
\begin{equation} 
  Q_{10} = \left( \frac{R_{2}}{R_{1}} \right)^{ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
  (\#eq:q10)
\end{equation} 
$$ Here, $R_{1}$ and $R_{2}$ are mean physiological rates and $T_{1}$ and $T_{2}$ are the temperatures that these rates are measured. Log transformation of equation \@ref(eq:q10) leads to the following log transformed $Q_{10}$:

$$
\begin{equation} 
  lnRR_{Q_{10}} = ln\left( \frac{R_{2}}{R_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
(\#eq:lnq10)
\end{equation} 
$$ Equation \@ref(eq:lnq10) is essentially a temperature corrected equivalent to lnRR when the numerator and denomenator are measured at different temperatures. This allows one to compare the mean of two temperature treatments directly regardless of the temperatures that these groups have been measured. Here, we will refer to this as the log $Q_{10}$ response ratio, $lnRR_{Q_{10}}$. Notably, by taking the natural logarithm, there are two statistical advantages: 1) 0 becomes biological meaningful as 0 means two rates are the exactly the same and 'statistically significantly different from 0' means that two rates are significantly different to each other and 2) $lnRR_{Q_{10}}$ is more likely to satisfy the assumption of residual normality than $Q_{10}$ (see @Hedges1999). Also, the recognition of this equivalence means that we can easily calculate the sampling variance for equation \@ref(eq:lnq10) as follows:

$$
\begin{equation} 
  s_{lnRR_{Q_{10}}} = \left( \frac{SD_{2}^2}{R^2_{2}N_{2}} + \frac{SD_{1}^2}{R^2_{1}N_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
  (\#eq:Vlnq10)
\end{equation} 
$$

#### *New Effect Size Metrics that Compare Changes in Variability Across Treatments*

@Nakagawa2015 recently proposed analogous effect size estimates to *lnRR* that allow for comparisons of changes in variance between two groups, the log variance ratio (*lnVR*) and the log coefficient of variation (*lnCVR*). Like *lnRR*, *lnVR* and *lnCVR* are ratios that describe the relative difference in trait variablity between two groups. We refer readers to @Nakagawa2015 for the equations describing *lnVR* and *lnCVR*, but these can easily be extended to their $Q_{10}$ analogues (and associated sampling variances) as follows:

$$
\begin{equation} 
lnVR_{Q_{10}} = ln\left( \frac{SD_{2}}{SD_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
(\#eq:lnq10VR)
\end{equation} 
$$ $$
\begin{equation} 
s_{lnVR_{Q_{10}}} = \left( \frac{1}{2\left( N_{2}-1 \right)} + \frac{1}{2\left(N_{1} - 1\right)} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
(\#eq:slnq10VR)
\end{equation} 
$$

Equations \@ref(eq:lnq10VR) and \@ref(eq:slnq10VR) describe the change in physiological rate variance (eqn \@ref(eq:lnq10VR)) across a 10°C temperature change along with its sampling variance (eqn \@ref(eq:slnq10VR)). While this is a useful metric, as discussed by @Nakagawa2015 there is often a strong mean-variance relationship that needs to be accounted for in analysing changes in variance. As such, we can calculate the coefficient of variation, which standardises changes in variance for changes in means as follows:

$$
\begin{equation} 
lnCVR_{Q_{10}} = ln\left( \frac{\text{CV}_{2}}{\text{CV}_{1}} \right){ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }
(\#eq:lnq10CVR)
\end{equation} 
$$ $$
\begin{equation} 
s_{lnCVR_{Q_{10}}} = \left[ \frac{(\text{SD}_{1})^2}{N_{1}({R}_{1})^2} + \frac{(\text{SD}_{2})^2}{N_{2} ({R}_{2})^2} + \frac{1}{2\left( N_{1}-1 \right)} + \frac{1}{2\left(N_{2} - 1\right)} \right]{ \left(\ \frac{10^{\circ}C}{T_{2}-T_{1}} \right) }^2
(\#eq:slnq10CVR)
\end{equation} 
$$

where $CV$ is the coefficient of variation defined as $SD / R$. Whether one choose to use SD or CV-based effect size depends on questions at hand (discussed in @Nakagawa2015 and Senior et al. 2020 <!--- sorry I have not added it to bib file - I will do this later once I go through --->)

## **Meta-Analytic Models to Control for and Investigate into Heterogeneous Effects Within and Across Studies**

<!--- As said before you want to play up the aspect of investigaing heterogeneity rather than controlling for it - controlling for it seems to mean you still want to know the overall effect after getting rid of every confunders. Whilst it is also waht we want, we really want to explain the hetero --->

It will not be possible to derive an effect size that completely controls for known heterogeneity at the effect size level all the time. This limitation is partly because treatment effects are not be simple linear or exponential functions of the mean response making the effect sizes we discuss above unsuitable [e.g., thermal performance curves; @Noble2018]. For example, dosages applied to treatments can follow quite complex non-linear relationships, with exponential or quadratic relationships in relation to some mean response. When this occurs we recommend comparative physiologists apply meta-regression approaches, controlling for these known treatment differences as a moderator. There are simple statisical approaches a meta-analyst can use to then make the overall effects interpretable [@Schielzeth2010]. Before diving into meta-regression models, we will first overview the meta-analytic approaches or steps commonly used.

#### *Multi-level Meta-Analysis when the Overall Effect is of Interest*

One of the main goals of meta-analysis is to obtain an overall meta-analytic mean effect size estimate and some form of uncertainty around that mean, like a 95% confidence interval (CI). Including a prediction interval (PI) can also be quite useful because, in the context of a meta-analysis, it provides the range of effect size values we might expect to observe if a new study were to be conducted [@Nakagawa2021c]. Estimating an overall mean along with 95% CI (or PI) is generally done with a simple random effects or multilevel meta-analytic model that accounts for effect size sampling error [@Nakagawa2017; @NakagawaSantos2012]. More often than not, meta-analyses in comparative physiology will have complex hierarchical structure. For example, multiple effect sizes may be taken from a single study on different traits measured on the same individuals, or come from many different species that share an evolutionary history [@Noble2017]. As such, it is far more likely that comparative physiologists will need to apply a multi-level meta-analytic model to control for the various sources of non-independence [@NakagawaSantos2012; @Song2021; @Nakagawa2021; @Noble2017]. Multilevel models are also extremely useful in describing the various sources contributing to effect size variance. For the purpose of our discussion, we will assume the meta-analyst plans to use more advanced multilevel models. Having said that, it maybe sufficient for many meta-analyses to make use of traditional random effects models. We refer readers to excellent reviews and books that overview the distinctions between common (fixed) effect, random effect and multilevel models and when one might want (or not want) to apply these models [@Nakagawa2017; @NakagawaSantos2012; @Koricheva2013; @Borenstein2009; @Schmid2021].

Assume that we are interested in understanding the overall effect of, say, stress (i.e., corticosterone - CORT) on reproduction [e.g., see @Podmola2018] we might extract data from studies manipulating corticosterone in bird eggs. Of course, these studies are diverse, apply different dosages of corticosterone and are working across many species. If the studies are experimental in nature, than a standardised effect size, such as $ln_{RR}$ or Hedges' *d* might be applicable. We then might fit an overall, multilevel meta-analytic model as follows: $$
\begin{equation} 
Y_{i} = \beta_{0} + a_{k[i]} + sp_{k[i]} + s_{j[i]} + e_{i} + m_{i}\
(\#eq:MLMM)
\end{equation} 
$$ $$
\begin{equation} 
a_{k[i]}\sim N(0, \sigma^{2}_aA),\
sp_{k[i]}\sim N(0, \sigma^{2}_{sp}I),\
s_{j[i]}\sim N(0, \sigma^{2}_{sp}I),\
e_{i}\sim N(0, \sigma^{2}_eI),\
m_{i} \sim N(0, v_{i})\
(\#eq:dist)
\end{equation} 
$$

<!--- you can use ni instead of spi??? - also refer to this paper in this seciton which really justify or even nessates people using this rather complext model and having a and sp are rather crtical! - https://ecoevorxiv.org/su4zv/--->

where $Y_{i}$ is the *i*th effect size (*i* = 1 ... $N_{ES}$, the number of effect sizes), $\beta_{0}$ is the overall meta-analytic mean, $a_{k[i]}$ is the phylogenetic effect (a random effect) for species *k* applied to effect size *i*. Phylogenetic effects are assumed to be normally distributed deviates sampled from a distribution with a mean of 0 and variance $\sigma^{2}_aA$ [i.e., *N*$(0, \sigma^{2}_aA)$], where A is a phylogenetic correlation matrix derived from a phylogenetic tree, $sp_{k[i]}$ is the species-specific effect for the *k*th species applied to effect size *i*, $s_{j[i]}$ is the study-specific effect for the *j*th study applied to effect size *i*, $e_{i}$ is the effect size-specific effect (or within study effect, or residuals) for the *i*th effect size, and $m_{i}$ is the sampling variance effect around the *i*th effect as a result of variation in precision among each effect sizes, $v_{i}$, which is known. This model is a 'weighted' multilevel meta-analysis model because the estimation of *u* is weighted by the inverse sampling variance of each effect size. The fact that we can obtain an overall estimate of sampling variance (i.e., $\sigma^2_{m}$) is very important because it provides a way for us to understand just how much variation in effects are the result of sampling differences across studies versus other, potentially important sources of variation such as differences in the methods applied, species differences or even real biological effects that impact the effect size a study is likely to observe [@Nakagawa2017; @NakagawaSantos2012]. This is formalised in what we call 'heterogeneity analysis' which we describe below. Importantly, this model provides us with the answer to our overall question: Across all the species and studies manipulating egg corticosterone, how much of an overall effect does corticosterone have on reproduction relative to control conditions? <!--- don't we need to mention in the case of VCV for mi--->

Of course, the overall effect does not provide us a way to interpret its magnitude in the context of how much CORT has been applied in the treatment. The most we can say about it, is that the effect is small, moderator of large out of the sample of effect sizes extracted from this set of studies. In the next section, we show how to 'correct' the overall effect size so that its interpretation is more meaningful.

#### *Multilevel Meta-regression to Understand Variation and Improve Interpretation*

Heterogeneity among effects that result from known differences in how treatments are applied can also be dealt with using multilevel meta-regression models. Multilevel meta-regression models include all the same random effects that we have already discussed, but they also include fixed effects (i.e., predictors or moderators) that attempt to explain changes in the mean effect size. In otherwords, we can modify our multilevel meta-analytic model to turn it into a multilevel meta-regression model as follows:

$$ Y_{i} = \beta_{0} + \sum \beta_{j}x_{j} + a_{k[i]} + sp_{k[i]} + s_{j[i]} + e_{i} + m_{i} $$ where $\sum \beta_{j}x_{j}$ is simply the sum of all effects for all moderators (fixed effects), $x_{j}$. All other notation is the same as described above. Importantly, the variable $x_{j}$ could be any moderator collected from studies, including dosage, temperature, salinity, or pH. Some moderators will be at the level of the study (e.g., was the study experimental or observational in nature) but others could be taken at the effect size level (i.e., was the effect derived using a tempearture of 23°C or 35°C). By including these moderators in a meta-regression model, the interpretation of the overall meta-analytic mean, $\beta_{0}$, changes. For example, assume we collected the change in CORT dosage relative to a control group (i.e., in uL / g) for our meta-analysis on egg CORT effects on bird reproduction. If the untransformed dosage, $x_{j}$, is included in the model then the overall meta-analystic mean estimate, $\beta_{0}$, would be the mean effect of CORT relative to a control treatment, on reproduction when the dosage difference between the treatment and control group is 0 uL /g. Of course, this interpretation does not make much sense. We can, however, apply some simple transformations to make the overall mean estimate more intuitive to understand with the added benefit that it can help improve the interpretation of non-linear parameters estimated by the model [@Schielzeth2010; @GelmanHill2007]. One simple transformation would be to centre the variable $x_{j}$, by subtracting each value from the mean to create a new input variable (i.e., $c_{i}=x_{i}-\bar{x}$, where the subscript *i* denotes each individual effect size dosage). The new variable, $c_{j}$, is now centered on the mean and can replace $x_{j}$. When the model is refit with $c_{j}$ the estimated $\beta_{0}$ can now be more intuitivly interpreted as the overall mean effect of CORT on reproduction at an average dosage difference between treatment and control groups. Importantly, by centering, we retain the original units (uL / g) of the variable, which can be very useful to allow for a more general interpretation of the effect magnitude. Of course, if one is not interested in the magnitude of effect, we could also change our question and simply model mean reproductive output as a function of dosage (often what are referred to as 'arm-based' meta-analytic models). Here, one can model the mean for each group (i.e., control and treatment group), and acount for the known sampling variance associated with each group mean. Dosage, or other continuous variables like temperature, can then be modelled (accounting for sampling variance) to understand how mean reproductive output changes. It is important to recognise that all means need to be comparable and that these models are more complex to fit. While they can be equivalent to 'contrast-based' models, they require complex interactions to be estimated [@Nakagawa2015].

Centering moderator variables, like temperature and dosage, not only provides a way to make the overall mean effect more interpretable in the face of treatment heterogenity, but it also allows more complex relationships to be fit without compromising interpretations, including non-linear parameters and interactions (e.g., polynomials). In other words, linear / main effects can still be interpreted normally in the presence of non-linear / interactions fit in the model [@Schielzeth2010; @GelmanHill2007]. This can be particularly powerful when more complex, nonlinear, dosage relationships exist. We provide examples in the online supplementary material (with R code) to show readers how teh models we describe above can be fit, and how, meta-analytic means can be adjusted for treatment heterogenity.

#### *Heterogenity Analysis: How Much Effect Size Variation Actually Exists?*

Quantifying effect size heterogeneity (i.e., effect size variance) provides a pathway to explore hypotheses thought to explain why effects might vary within and across studies. Effect size heterogeneity can be thought of as the percentage of variance between effect sizes that cannot be attributable to sampling error (often as called 'total' heterogeneity) [@Higgins2002]. More formally, we can quantify heterogeneity using the multilevel meta-analytic model we describe above as:

$$ 
I^2_{total} = \frac{\sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual}}{\sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual} +\sigma^2_{m}}
$$

<!--- be consistent with sigma notations above - actuall total - m may sometimes not work - so best adding up all like this, which was the core finding in Nakagawa & Schielzeth's R2 --->

where $\sigma^2_{total} = \sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual} +\sigma^2_{m}$ is the total effect size variance (using the above multilevel model) and $\sigma^2_{m}$ is 'typical' the sampling error variance. <!--- you need more than that sigma_m is a typical one - you may need to provide a formula here --->. Quantifying heterogeneity provides a formal way to assess how much variation is the result of 'true' biological or methodological differences as opposed to simply chance (i.e., sampling variance) [@NakagawaSantos2012; @Higgins2002]. We can only conduct such analyses if we have an estimate of the sampling variance for a given effect measure because only then can we quantify how much variation is the result of sampling variability. @Senior2016 have shown that total heterogeneity is often extremely high in ecological and evolutionary meta-analysis (\~91%), suggesting that effect measures vary widely across and within studies. Some of the variation is clearly the result of working with different species or strains, different traits and different outcome measures, however, some of this variation also arises because of methodological differences across studies that might be related to the dosages or temperatures used for treatments. Formally quantifying and presenting heterogeneity estimates from meta-analytic models can provide a way to understand the major drivers of effect size variation by allowing researchers to produce various $I^2$ estimates which can be formally compared within and across studies. For example, we could fit the multilevel model described above to understand what proportion of variation is the result of, say, 'study' effects following @NakagawaSantos2012:

$$ I^2_{study} = \frac{\sigma^2_{study}}{\sigma^2_{total}} $$ where $\sigma^2_{study}$ is the study specific variance, and all other variance components are for phylogeny, species, residual (effect size level) and m (total measurement/sampling error). Many readers will notice the similarities between $I^2$ and $R^2$ [@Nakgawa2013; @Nakagawa2010]. If we want to formally assess how much effect size variation is the result of methodological factors, such as temperature and dosage differences applied across studies, then we could fit our multilevel meta-regression model, including dosage or temperature (linear, quadratic or even more complex fits), and estimate how much variation is explained by these fixed effects following @Nakgawa2013.

$$ R^2_{marginal} = \frac{\sigma^2_{fixed}}{\sigma^2_{fixed} + \sigma^2_{study} + \sigma^2_{phylogeny} + \sigma^2_{species} + \sigma^2_{residual}} $$

Note that this formula does not include $\sigma^2_{m}$ as sampling error variance is assumed to be known in meta-analysis as explained above <!--- if you have not done so, please do --->. The $R^2_{marginal}$ provides a formal means to assess just how much variation in effects is teh result of known heterogeneity at the effect size level <!--- note what level's variation it accounts for depends on the type of a fixed effect or moderator --->, driven by continuous moderators such as dosage, temperature, salinity, and even pH. $I^2_{hetero}$ maybe a useful measure to provide to help readers understand just how much effect variation can result from the specific treatment application. <!--- you can also say metaAider and orcharRd packages have functions to cacualte this already via metafor - implemention is important --->

## **Conclusions and Future Directions**

## **References**
